PLANETESIMAL ACCRETION AT SHORT ORBITAL PERIODS

### 2022-07-21 main.tex
TRQ edits through figure 3.

In section \ref{sec:theory} we provide an overview of the theory behind planetesimal accretion and show that assumptions used to derive the well-known modes of growth are only valid at sufficiently long orbital periods. We then motivate the need for N-body simulations to study this problem and describe the code used, along with how our initial conditions were constructed in section \ref{sec:methods}. In section \ref{sec:narrow}, we present a parameter study of planetesimal accretion using a series of simulations of narrow annuli at various orbital periods. We also present a set of simulations starting with a much wider planetesimal disk and demonstrate that a transition between accretion modes occurs at moderately small ($\simeq$ 50d) orbital periods. Next, we assess the impact of simplifications made to our collision model on this result in section \ref{sec:assump}. In section \ref{sec:discuss}, we discuss the implications of this multimodal accretion behavior throughout the disk for planet formation models and conclude.

@@

We begin our analysis by considering a disk of equal planetesimals
with radius $r_{pl}$, mass $m_{pl}$ and surface density
$\Sigma_{pl}$. The collision rate in the vicinity of an orbit defined
by Keplerian frequency $\Omega$ can be written as $n \Gamma v$, where
$n = \Sigma_{pl} \Omega / 2 m_{pl} v$, $\Gamma$ is the effective
collision cross section and $v$ is the typical encounter velocity
between planetesimals.
% TRQ: mention that $v$ is related to the velocity dispersion of the
% planetesimals.
Assuming that every collision results in a perfect merger, the growth rate of a planetesimal is given by

\begin{equation}\label{eq:growth}
	\frac{dM}{dt} = \frac{\Sigma \Omega}{2 m_{pl}} \Gamma.
\end{equation}

@@

In the limit that $v_{esc} \gg v$, it can be shown that $M \propto M^{4/3}$, which implies a runaway scenario, in which growth accelerates with mass. This mode of growth was confirmed with N-body simulations by \citet{kokubo96} and appears necessary to construct protoplanets within the lifetime of a protoplanetary disk. Due to the velocity dependence of the gravitational focusing effect, it isn't clear how ubiquitous this mode of growth is. In particular, encounter velocities at short orbital periods will be rather large (because $v \sim v_{k}$) and the $v_{esc} \gg v$ condition may not always be satisfied. The effect that a dynamically hot disk has on runway growth will be examined in detail in section \ref{sec:narrow}.

An important feature is missing from the model described above, which
limits its applicability at late times. Encounters between
planetesimals that do not result in a collision play a crucial role
over long timescales. These close encounters act to convert energy
from Keplerian shear into random motion. Over time, this gravitational
stirring effect will raise the typical encounter velocities between
bodies and diminish the effectiveness of gravitational focusing. With
a spectrum of masses, these velocity differences become even more
pronounced because as system evolves towards a state of quasi-equilibrium where $v \sim m^{1/2}$. For a system of equal mass bodies in which encounters are driven by random motions rather than Keplerian shear (dispersion dominated), the timescale for gravitational stirring is described by the two-body relaxation time \citep{ida93}

\begin{equation}\label{eq:relax}
	t_{relax} = \frac{v^3}{n \pi G^2 m_{pl} \ln \Lambda},
\end{equation}

\noindent where $\mathrm{ln} \Lambda$ is the Coulomb logarithm,
typically taken to be $\approx 10$ for a planetesimal disk. Despite
the fact that the behavior of gravitational stirring is well-described
by the two-body formalism, \citep{ida93} found that the stirring in a planetesimal disk is actually driven by close encounters. As we will show in section \ref{sec:narrow}, gravitational stirring effectively shuts off when the Hill sphere of a body becomes comparable to its physical size. In this case, close encounters tend to result in collisions, and the main pathway for energy exchange between planetesimals is unable to operate.

As mentioned above, the system will tend toward a state of energy
equipartition for a non-uniform mass distribution. \citet{kokubo98}
showed that runaway growth is actually self-limiting. As the runaway
bodies grow larger, they become more effective at heating the
remaining planetesimals, which diminishes the effectiveness of
gravitational focusing and throttles the growth rate. Around the time
that the massive runaway bodies make up about 50 percent of the solid
surface density (citation), a period of less vigorous ``oligarchic'' growth commences, in which the few largest bodies continue to accrete planetesimals at similar rates.

@@

In figure \ref{fig:timescales}, we show ratio between the relaxation and collision timescale for a population of equal-mass planetesimals as a function of orbital period. Here, the encounter velocity is described by $v = \left( 5/8 \left< e^2 \right>^{1/2} + \left< i^2 \right>^{1/2} \right)^{1/2} v_{k}$, where $\left< e^2 \right>^{1/2}$ and $\left< i^2 \right>^{1/2}$ are the eccentricity and inclination dispersions of the planetesimal disk and $v_{k}$ is the Keplerian velocity. For simplicity, we assume that $\left< e^2 \right>^{1/2} = 2\left< i^2 \right>{1/2}$ \citep{ida93a} and that the eccentricity dispersion is constant with orbital period. The eccentricity dispersion is described in units scaled by the Hill factor $\left( m_{pl}/ 3 M_{star} \right)^{1/3}$ such that $e_{h} = 1$ corresponds to the boundary between shear and dispersion dominated encounters. The horizontal dashed line indicates where $t_{relax} = t_{coll}$. The timescale criterion for oligarchic growth is only satisfied in regions where the disk is sufficiently dynamically cold and the orbital period is sufficiently long. In section \ref{sec:results} we will explore the behavior and outcome of planetesimal accretion in regions where this criterion is \textit{not} satisfied.

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/timescales.png}
    \caption{The ratio between the two-body relaxation and collision timescale for a population of equal-mass planetesimals with an internal density of 3 g cm$^{-3}$. Only in regions where $t_{relax} \ll t_{coll}$ can the velocity distribution respond to changes in the mass of the bodies such that oligarchic growth can operate. This condition is no longer satisfied for a dynamically hot disk at sufficiently short orbital periods.\label{fig:timescales}}
\end{center}
\end{figure}

@@

An insight into the expected behavior in this regime can be gained by defining the dimensionless parameter $\alpha$, which is the ratio between the physical size of a body and it's Hill radius

\begin{equation}\label{eq:alpha}
	\alpha = \frac{r_{pl}}{r_{h}} = \frac{1}{a} \left( \frac{9 M_{*}}{4 \pi \rho_{pl}} \right)^{1/3},
\end{equation}

\noindent where $a$ is the semimajor axis of the body and $\rho_{pl}$ is its bulk density. Assuming a fixed bulk density as bodies collide and grow, and that no large-scale migration occurs, the scaling of both $r_{pl}$ and $r_{h}$ with $m_{pl}^{1/3}$ means that $\alpha$ will be constant with time. For a composition of ice and rock, $\alpha$ is small for any populated region of the solar system ($\alpha \sim 10^{-2}$ for Earth and $\alpha \sim 10^{-4}$ in the Kuiper belt). As one moves close to the star $\alpha$ becomes larger than 1, which implies that the physical extent of a body exceeds its Hill sphere. (also mention that hill sphere size independent of velocity for small eccentricity)

@@

We also introduce a second dimensionless quantity, which relates the physical size of the bodies to the velocity state of the system

\begin{equation}\label{eq:beta}
	\beta = \frac{r_{pl}}{r_{g}}.
\end{equation}

\noindent where $r_{g} = G m_{pl} / v^{2}$ is the gravitational radius of a body. Encounters between bodies inside of a distance of $r_{g}$ result in significant deflections of their trajectories. It should be noted that the gravitational focusing enhancement factor $v^{2}/v_{esc}^{2}$ is equal to 1 for $\beta = 1$. In the case where $r_{g}$ is smaller than the size of a planetesimal, the gravitational focusing enhancement factor will be between 0 and 1.

Need a transition to next section here.

@@

We use the tree-based N-body code {\sc ChaNGa} \footnote{A public version of {\sc ChaNGa} can be downloaded from \url{http://www-hpcc.astro.washington.edu/tools/ChaNGa.html}} to model the gravitational and collisional evolution of planetesimals at short orbital periods (where $\alpha$ is large). {\sc ChaNGa} is written using the {\sc CHARM++} parallel programming language and has been shown to perform well on up to half a million processors \citep{menon15} and can follow the evolution of gravitationally interacting swarms of up to billions of particles. Using a modified Barnes-Hut tree with hexadecapole expansions of the moments to approximate forces, {\sc ChaNGa} integrates the equations of motion using a kick-drift-kick leapfrog scheme. For all of the simulations presented in this paper, we use a node opening criteria of $\theta_{BH}$ = 0.7. Additional information about the code is available in \citep{jetley08}.

Using the neighbor-finding algorithm in {\sc ChaNGa}, originally designed for SPH calculations, we have recently implemented a solid body collision module in the code.This work is largely based on the solid-body collision implementation in {\sc PKDGRAV}, which is described in \citet{richardson94} and \citet{richardson00}. To summarize, imminent collisions are detected during the `drift' phase by extrapolating positions of bodies forward in time, using the velocity calculated at the opening `kick'. For each body, any neighboring particles which fall within a search ball of radius $fill in later$ are tested for an imminent collision. In the case that a collision is detected, the particles are merged into a single larger body, which is given the center of mass position and velocity of the two children. Resolving a collision can produce another imminent collision, so collisions are handled one-by-one and another full collision check is run after the previous event is resolved. For a more detailed description of the collision module in {\sc ChaNGa}, see \citep{wallace19}. (mention grav stepping here as well, since its used for all simulations)

@@

We begin by presenting a parameter study of planetesimal accretion, motivated by two questions raised in section \ref{sec:theory}. 1. Does runaway growth still operate when the condition that $v \ll v_{esc}$ is not satisfied? 2. How does planetesimal accretion proceed when the planetesimals themselves occupy a significant fraction of their Hill spheres?

To answer these questions, we run a series of simulations in which a narrow annulus of planetesimals orbits a star. The values of $\alpha$ and $\beta$ are varied individually. 4000 planetesimals with individual masses of $5 \times 10^{23}$ g are placed with semimajor axes randomly drawn between 0.95 and 1.05 AU about a 1 $M_{\odot}$ star. The argument of perihelion $\omega$, longitude of ascending node $\Omega$, and mean anomaly M for each body is drawn from a uniform distribution $\in [0, 2 \pi)$. The inclination dispersion $\left< i^{2} \right>^{1/2}$ is chosen to be half the eccentricity dispersion $\left< e^{2} \right>^{1/2}$ \citep{ida93a}.

% Mention Rayleigh distribution
In the `nominal' case, we give the bodies a bulk density of 3 g cm$^{-3}$, and $\left< e^{2} \right>^{1/2} = 4 e_{h}$, where $e_{h} = e a/r_{h}$ (define this in the previous section instead) which corresponds to $\alpha = 3.6 \times 10^{-2}$ and $\beta = 3.4 \times 10^{-3}$. These parameters are chosen to match the initial conditions of \citet{kokubo98}, which gave rise to oligarchic growth. To vary the value of $\alpha$, we alter the bulk density of the particles. In the high-$\alpha$ case, the bulk density is reduced by a factor of $\sim$ 7100, which produces $\alpha = 1$. To vary $\beta$, the eccentricity dispersion is increased. For the high-$\beta$ case, $\left< e^{2} \right>^{1/2}$ is increased to $1500 e_{h}$, which corresponds to $\beta = 15,000$.

In all cases, the simulations are evolved with a base timestep of 1.7 days, which corresponds to 3\% of an orbital dynamical time $\sqrt{a^3/G M_{*}}$. Due to the vastly differing growth timescales in each case, a simulation is stopped when the growth of the most massive body flattens out. In figure \ref{fig:alpha_beta}, we show the a-e distribution of bodies in the initial (blue) and final (orange) snapshots from each of the 4 simulations. The size of the points indicates the relative masses of the bodies. Only in the case of small $\alpha$ does a residual population of dynamically hot planetesimals develop. The lack of high eccentricity planetesimals in the large $\alpha$ case suggests that most encounters instead result in accretion. In the case of large $\beta$, the protoplanets and remaining planetesimals end up in a dynamically cool state, compared to the initial conditions. (Are the inelastic collisions playing a significant role here?)

@@

To investigate whether any of these planetesimal disks underwent runaway growth, we examine the time evolution of the maximum and mean masses in each case. The ratio $m_{max}/\left< m \right>$ is plotted in figure \ref{fig:alpha_beta_evo}. On this plot, a positive slope indicates that the growth rate is accelerating with mass, which is evident in all four cases. Even with a large value of $\beta$, which means that the effective collision cross section is very near the geometric value, runaway growth still appears to operate. This seems to imply that, as bodies collide and grow, the relative increase in their gravitational focusing factors, rather than its absolute size, is what drives the system towards runaway growth. Although larger encounter velocities will lengthen the growth timescales, this mode of growth appears to be inevitable, so long as gravity is the dominant force in the system.

@@

Motivated by the apparent dependence of accretion modes on $\alpha$, we next investigate whether this highly efficient, non-oligarchic growth mode should be expected to operate in a typical planet-forming disk. Given the dearth of short period terrestrial planets observed around M stars (citation), we model a series of wide planetesimal disks, which span from 1 to 100 days in orbital period orbiting a late-type M star of mass 0.08 $M_{\odot}$. For a population of planetesimals with a bulk density of 3 g cm$^{-3}$, this orbital period range corresponds to $\alpha \in (0.7, 0.05)$. By simultaneously modeling a broad range of orbital periods, we can determine the critical value of $\alpha$ that divides these two modes of accretion, and also explore how the oligarchic/non-oligarchic accretion boundary affects the resulting distribution of protoplanets.

Four wide disk simulations are run in total (see table \ref{tab:sims}). In each case, the solid surface density follows a power law profile

\begin{equation}
	\Sigma(r) = \left( \frac{M_{*}}{M_{\odot}} \right) A \Sigma_{mmsn} \left( \frac{r}{1 AU} \right)^{-\alpha},
\end{equation}

\noindent where $M_{*}$ is the mass of the central star, $\Sigma_{MMSN} = 10$ g cm$^{-2}$ is surface density of the minimum-mass solar nebula \citep{hayashi81} at 1 AU, and $A$ is an enhancement factor. In the first case (fdHi), we model a disk that follows a MMSN power law slope, with the overall normalization enhanced by a factor of 100. This choice of parameters for the solid surface density profile appears necessary in order to reproduce many observed short period terrestrial worlds in-situ \citep{hansen12}. Additionally, we vary the power law index (fdHiShallow, fdHiSteep) and overall normalization (fdLo) of the solids profile.

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/alpha_beta_mass.png}
    \caption{The final state of the mass distribution for the simulations described in section \ref{sec:narrow}. For small $\alpha$, a few embryos form alongside a power law tail of planetesimals. For larger values of $\alpha$, the mass distribution takes on a more unimodal form. As in the previous figure, the initial choice of $\beta$ does not appear to have any meaningful impact on the end result.\label{fig:alpha_beta_mass}}
\end{center}
\end{figure}

>>>>

In section \ref{sec:theory} we provide an overview of the theory
behind planetesimal accretion and show that assumptions used to derive
the well-known modes of growth are only valid at sufficiently long
orbital periods. We then motivate the need for N-body simulations to
study this problem and describe the code used, along with how our
initial conditions were constructed in section \ref{sec:methods}. In
section \ref{sec:narrow}, we present a parameter study of planetesimal
accretion using a series of simulations of narrow annuli at various
orbital periods. In section \ref{sec:fulldisk} we present a set of simulations starting with a much wider planetesimal disk and demonstrate that a transition between accretion modes occurs at moderately small ($\simeq$ 50d) orbital periods. Next, we assess the impact of simplifications made to our collision model on this result in section \ref{sec:assump}. In section \ref{sec:discuss}, we discuss the implications of this multimodal accretion behavior throughout the disk for planet formation models and conclude.

@@

We begin our analysis by considering a disk of equal planetesimals
with radius $r_{pl}$, mass $m_{pl}$ and surface density
$\Sigma_{pl}$. The collision rate in the vicinity of an orbit defined
by Keplerian frequency $\Omega$ can be written as $n \Gamma v$, where
$n = \Sigma_{pl} \Omega / 2 m_{pl} v$, $\Gamma$ is the effective
collision cross section and $v$ is the typical encounter velocity
between planetesimals.
% TRQ: mention that $v$ is related to the velocity dispersion of the
% planetesimals.  Also: mention that the scale height of the disk is
% assumed to be \Omega/(2v)
Assuming that every collision results in a perfect merger, the growth rate of a planetesimal is given by
\begin{equation}\label{eq:growth}
	\frac{dM}{dt} = \frac{\Sigma \Omega}{2 m_{pl}} \Gamma.
\end{equation}

@@

In the limit that $v_{esc} \gg v$, it can be shown that $M \propto
M^{4/3}$, which implies a runaway scenario, in which growth
accelerates with mass. This mode of growth was confirmed with N-body
simulations by \citet{kokubo96} and appears necessary to construct
protoplanets within the lifetime of a protoplanetary disk. Due to the
velocity dependence of the gravitational focusing effect, it is not clear how ubiquitous this mode of growth is. In particular, encounter velocities at short orbital periods will be rather large (because $v \sim v_{k}$) and the $v_{esc} \gg v$ condition may not always be satisfied. The effect that a dynamically hot disk has on runway growth will be examined in detail in section \ref{sec:narrow}.

An important feature is missing from the model described above, which
limits its applicability at late times. Encounters between
planetesimals that do not result in a collision play a crucial role
over long timescales. These close encounters act to convert energy
from Keplerian shear into random motion. Over time, this gravitational
stirring effect will raise the typical encounter velocities between
bodies and diminish the effectiveness of gravitational focusing. With
a spectrum of masses, these velocity differences become even more
pronounced because as the system evolves, it moves to a state of
energy equipartition where $v \sim m^{1/2}$. For a system of equal mass bodies in which encounters are driven by random motions rather than Keplerian shear (dispersion dominated), the timescale for gravitational stirring is described by the two-body relaxation time \citep{ida93}
\begin{equation}\label{eq:relax}
	t_{relax} = \frac{v^3}{4 \pi n G^2 {m_{pl}}^2 \ln \Lambda},
\end{equation}
%TRQ: changed this to match Ida & Makino 1993.
where $\ln \Lambda$ is the Coulomb logarithm,
typically taken to be $\approx 10$ for a planetesimal disk. Despite
the fact that the behavior of gravitational stirring is well-described
by the two-body formalism, \citep{ida93} found that the stirring in a planetesimal disk is actually driven by close encounters. As we will show in section \ref{sec:narrow}, gravitational stirring effectively shuts off when the Hill sphere of a body becomes comparable to its physical size. In this case, close encounters tend to result in collisions, and the main pathway for energy exchange between planetesimals is unable to operate.

As mentioned above, the system will tend toward a state of energy
equipartition for a non-uniform mass distribution. \citet{kokubo98}
showed that runaway growth is actually self-limiting. As the runaway
bodies grow larger, they become more effective at heating the
remaining planetesimals, which diminishes the effectiveness of
gravitational focusing and throttles the growth rate. Around the time
that the massive runaway bodies make up about 50 percent of the solid
surface density
%TRQ: citation needed.
(citation), a period of less vigorous ``oligarchic'' growth commences, in which the few largest bodies continue to accrete planetesimals at similar rates.

@@

In figure \ref{fig:timescales}, we show ratio between the relaxation
and collision timescale for a population of equal-mass planetesimals
as a function of orbital period. Here, the encounter velocity is
described by
$$
v = \left( 5/8 \left< e^2 \right>^{1/2} + \left< i^2 \right>^{1/2}
\right)^{1/2} v_{k},
$$
(citation) where $\left< e^2 \right>^{1/2}$ and $\left< i^2
\right>^{1/2}$ are the eccentricity and inclination dispersions of the
planetesimal disk and $v_{k}$ is the Keplerian velocity. For
simplicity, we assume that $\left< e^2 \right>^{1/2} = 2\left< i^2
\right>^{1/2}$ \citep{ida93a} and that the eccentricity dispersion is
constant with orbital period. The eccentricity dispersion is described
in units scaled by the Hill factor $\left( m_{pl}/ 3 M_{\star}
\right)^{1/3}$ such that $e_{h} = 1$ corresponds to the boundary
between shear and dispersion dominated encounters. The horizontal
dashed line indicates where $t_{relax} = t_{coll}$. The timescale
criterion for oligarchic growth is only satisfied in regions where the
disk is sufficiently dynamically cold and the orbital period is
sufficiently long. In sections \ref{sec:narrow} and \ref{sec:fulldisk}
we will explore the behavior and outcome of planetesimal accretion in regions where this criterion is \textit{not} satisfied.

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/timescales.png}
    \caption{The ratio between the two-body relaxation and collision
      timescale for a population of equal-mass planetesimals with an
      internal density of 3 g cm$^{-3}$ and an eccentricity dispersion
      characterized by $e_h$, where $e_h = 1$ is the boundary between
      shear and dispersion dominated encounters. Only in regions where $t_{relax} \ll t_{coll}$ can the velocity distribution respond to changes in the mass of the bodies such that oligarchic growth can operate. This condition is no longer satisfied for a dynamically hot disk at sufficiently short orbital periods.\label{fig:timescales}}
\end{center}
\end{figure}

@@

An insight into the expected behavior in this regime can be gained by
defining the dimensionless parameter $\alpha$, which is the ratio
between the physical size of a body and it's Hill radius, $r_{h}$:
\begin{equation}\label{eq:alpha}
	\alpha = \frac{r_{pl}}{r_{h}} = \frac{1}{a} \left( \frac{9 M_{\star}}{4 \pi \rho_{pl}} \right)^{1/3},
\end{equation}
where $a$ is the semimajor axis of the body and $\rho_{pl}$ is its bulk density. Assuming a fixed bulk density as bodies collide and grow, and that no large-scale migration occurs, the scaling of both $r_{pl}$ and $r_{h}$ with $m_{pl}^{1/3}$ means that $\alpha$ will be constant with time. For a composition of ice and rock, $\alpha$ is small for any populated region of the solar system ($\alpha \sim 10^{-2}$ for Earth and $\alpha \sim 10^{-4}$ in the Kuiper belt). As one moves close to the star $\alpha$ becomes larger than 1, which implies that the physical extent of a body exceeds its Hill sphere. (also mention that hill sphere size independent of velocity for small eccentricity)

@@

We also introduce a second dimensionless quantity, which relates the physical size of the bodies to the velocity state of the system
\begin{equation}\label{eq:beta}
	\beta = \frac{r_{pl}}{r_{g}}.
\end{equation}
where $r_{g} = G m_{pl} / v^{2}$ is the gravitational radius of a body. Encounters between bodies inside of a distance of $r_{g}$ result in significant deflections of their trajectories. It should be noted that the gravitational focusing enhancement factor $v^{2}/v_{esc}^{2}$ is equal to 1 for $\beta = 1$. In the case where $r_{g}$ is smaller than the size of a planetesimal, the gravitational focusing enhancement factor will be between 0 and 1.

These scaling considerations motivate the range of parameters in the
following numerical experiments as explore the different regimes.

% Need a transition to next section here.

@@

We use the tree-based N-body code {\sc ChaNGa}\footnote{A public version of {\sc ChaNGa} can be downloaded from \url{http://www-hpcc.astro.washington.edu/tools/ChaNGa.html}} to model the gravitational and collisional evolution of planetesimals at short orbital periods (where $\alpha$ is large). {\sc ChaNGa} is written using the {\sc CHARM++} parallel programming language and has been shown to perform well on up to half a million processors \citep{menon15} and can follow the evolution of gravitationally interacting swarms of up to billions of particles. Using a modified Barnes-Hut tree with hexadecapole expansions of the moments to approximate forces, {\sc ChaNGa} integrates the equations of motion using a kick-drift-kick leapfrog scheme. For all of the simulations presented in this paper, we use a node opening criteria of $\theta_{BH}$ = 0.7. Additional information about the code is available in \citep{jetley08,menon15}.

Using the neighbor-finding algorithm in {\sc ChaNGa}, originally
designed for SPH calculations, we have recently implemented a solid
body collision module in the code.  This work is largely based on the
solid-body collision implementation in {\sc PKDGRAV}, which is
described in \citet{richardson94} and \citet{richardson00}. To
summarize, imminent collisions are detected during the ``drift'' phase
by extrapolating positions of bodies forward in time, using the
velocity calculated at the opening ``kick''. For each body, any
neighboring particles which fall within a search ball of radius XXX
$fill in later$ are tested for an imminent collision. In the case that
a collision is detected, the particles are merged into a single larger
body, which is given the center of mass position and velocity of the
two children. Resolving a collision can produce another imminent
collision, so collisions are handled one-by-one and another full
collision check is run after the previous event is resolved. For a
more detailed description of the collision module in {\sc ChaNGa}, see
\citep{wallace19}.
Particles are advanced in time on individual timesteps chosen as a
power of two of a base timestep.  The timestep for an individual
particle is based on an estimate of the gravitational dynamical time
determined by the minimum of $\sqrt{d_{node}^3/(G(M_{node} + m_{pl}))}$
across all nodes in the tree that are accepted by the Barnes-Hut
opening criterion.  Here $d_{node}$ is the distance from the
planetesimal to the center of mass of the tree node and $M_{node}$ is
the total mass of the tree node.  For nearby particles $M_{node}$ is
replaced with the mass of the nearby particle.

@@

We begin by presenting a parameter study of planetesimal accretion,
motivated by two questions raised in section \ref{sec:theory}. 1) Does
runaway growth still operate when the condition that $v \ll v_{esc}$
is not satisfied? 2) How does planetesimal accretion proceed when the planetesimals themselves occupy a significant fraction of their Hill spheres?

To answer these questions, we run a series of simulations in which a
narrow annulus of planetesimals orbits a star. The values of $\alpha$
and $\beta$ are varied individually. 4000 planetesimals with
individual masses of $5 \times 10^{23}$ g are placed with semimajor
axes randomly drawn between 0.95 and 1.05 AU about a 1 $M_{\odot}$
star. The argument of perihelion $\omega$, longitude of ascending node
$\Omega$, and mean anomaly M for each body is drawn from a uniform
distribution $\in [0, 2 \pi)$. The inclination dispersion $\left<
i^{2} \right>^{1/2}$ is chosen to be half the eccentricity
dispersion $\left< e^{2} \right>^{1/2}$ \citep{ida93a}.
% Mention Rayleigh distribution: TRQ: yes!

In the ``fiducial'' case, we give the bodies a bulk density of 3 g
cm$^{-3}$, and $\left< e^{2} \right>^{1/2} = 4 e_{h}$, which corresponds to $\alpha = 3.6 \times 10^{-2}$ and $\beta = 3.4 \times 10^{-3}$. These parameters are chosen to match the initial conditions of \citet{kokubo98}, which gave rise to oligarchic growth. To vary the value of $\alpha$, we alter the bulk density of the particles. In the high-$\alpha$ case, the bulk density is reduced by a factor of $\sim$ 7100, which produces $\alpha = 1$. To vary $\beta$, the eccentricity dispersion is increased. For the high-$\beta$ case, $\left< e^{2} \right>^{1/2}$ is increased to $1500 e_{h}$, which corresponds to $\beta = 15,000$.

In all cases, the simulations are evolved with a base timestep of 1.7
days, which corresponds to 3\% of an orbital dynamical time
$\sqrt{a^3/G M_{*}}$. Due to the vastly differing growth timescales in
each case, a simulation is stopped when the growth of the most massive
body flattens out. In figure \ref{fig:alpha_beta}, we show the a-e
distribution of bodies in the initial (blue) and final (orange)
snapshots from each of the 4 simulations. The size of the points
indicates the relative masses of the bodies. Only in the case of small
$\alpha$ does a residual population of dynamically hot planetesimals
develop. The lack of high eccentricity planetesimals in the large
$\alpha$ case suggests that most encounters result in accretion
instead of scattering. In the case of large $\beta$, the protoplanets
and remaining planetesimals end up in a dynamically cool state,
compared to the initial conditions.
%TRQ albeit for small alpha, only slightly cooler.
(Are the inelastic collisions playing a significant role here?)

@@

To investigate whether any of these planetesimal disks underwent
runaway growth, we examine the time evolution of the maximum and mean
masses in each case. The ratio $m_{max}/\left< m \right>$ is plotted
in figure \ref{fig:alpha_beta_evo}. On this plot, a positive slope
indicates that the growth rate is accelerating with mass, which is
evident at various times in all four cases. Even with a large value of
$\beta$, which means that the effective collision cross section is
very near the geometric value, runaway growth still appears to
operate. This seems to imply that, as bodies collide and grow, the
relative increase in their gravitational focusing factors, rather than
its absolute size, is what drives the system towards runaway
growth. Although larger encounter velocities will lengthen the growth
timescales, this mode of growth appears to be inevitable, so long as
gravity is the dominant force in the system.
% TRQ: also mention the negative slope

@@

Motivated by the apparent dependence of accretion modes on $\alpha$,
we next investigate whether this highly efficient, non-oligarchic
growth mode should be expected to operate in a typical planet-forming
disk. Given the dearth (TRQ: why is this a reason for a wide disk?) of short period terrestrial planets observed around M stars (citation), we model a series of wide planetesimal disks, which span from 1 to 100 days in orbital period orbiting a late-type M star of mass 0.08 $M_{\odot}$. For a population of planetesimals with a bulk density of 3 g cm$^{-3}$, this orbital period range corresponds to $\alpha \in (0.7, 0.05)$. By simultaneously modeling a broad range of orbital periods, we can determine the critical value of $\alpha$ that divides these two modes of accretion, and also explore how the oligarchic/non-oligarchic accretion boundary affects the resulting distribution of protoplanets.

Four wide disk simulations are run in total (see table \ref{tab:sims}). In each case, the solid surface density follows a power law profile
\begin{equation}
	\Sigma(r) = \left( \frac{M_{*}}{M_{\odot}} \right) A \Sigma_{MMSN} \left( \frac{r}{1 AU} \right)^{-\alpha},
\end{equation}
%TRQ: should we use a different greek letter for the power law index?
where $M_{*}$ is the mass of the central star, $\Sigma_{MMSN} = 10$ g cm$^{-2}$ is surface density of the minimum-mass solar nebula \citep{hayashi81} at 1 AU, and $A$ is an enhancement factor. In the first case (fdHi), we model a disk that follows a MMSN power law slope, with the overall normalization enhanced by a factor of 100. This choice of parameters for the solid surface density profile appears necessary in order to reproduce many observed short period terrestrial worlds in-situ \citep{hansen12}. Additionally, we vary the power law index (fdHiShallow, fdHiSteep) and overall normalization (fdLo) of the solids profile.

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/alpha_beta_mass.png}
    \caption{The final state of the mass distribution for the
      simulations described in section \ref{sec:narrow}. For small
      $\alpha$, a few embryos form alongside a power law tail of
      planetesimals. For larger values of $\alpha$, the mass
      distribution takes on a more unimodal (TRQ: uniform?) form. As in the previous figure, the initial choice of $\beta$ does not appear to have any meaningful impact on the end result.\label{fig:alpha_beta_mass}}
\end{center}
\end{figure}
